A general Nullstellensatz for generalized spaces

### 2019-03-17 paper-qcoh.tex
Slight strengthening
typo

Theorem~\ref{thm:nullstellensatz} cannot be strengthened to arbitrary
first-order (or first-order$^\star$ or even just coherent) formulas in place of
geometric$^\star$ sequents. For instance, in the case that~$\TT$ is the theory
of local rings, the generic model~$U_\TT$ validates the coherent
formula~$\speak{any element which is not zero is invertible}$, but~$\TT/U_\TT$
does not prove this fact, as it is for instance not validated by the polynomial
algebra~$U_\TT[X]$. However, Theorem~\ref{thm:nullstellensatz} still plays an
important role in understanding first-order formulas:

\begin{thm}\label{thm:characterization}
Let~$\TT$ be a geometric theory. Let~$\alpha$ be a first-order
formula over the signature of~$\TT$. Then the following statements are
equivalent.
\begin{enumerate}
\item The formula~$\alpha$ holds for~$U_\TT$. \smallskip
\item The formula~$\alpha$ is provable in first-order intuitionistic logic
modulo the axioms of~$\TT$ and the additional axioms
\begin{equation}\label{eq:nullstellensatz}\tag{$\ddagger$}
  \speak{$\sigma$ holds for~$U_\TT$} \Longrightarrow \speak{$\ul{\TT}/U_\TT$
  proves$^\star$~$\sigma$}
\end{equation}
where~$\sigma$ ranges over the geometric formulas over the signature of~$\TT$.
\end{enumerate}
\end{thm}

>>>>

\begin{rem}In the case that the~$\Set[\TT']$ is a dense subtopos
of~$\Set[\TT]$, Theorem~\ref{thm:nullstellensatz-horn} can be strengthened to
allow~$\bot$ as consequent, since in this case~$(\nabla\bot) \Rightarrow \bot$.\end{rem}

Theorem~\ref{thm:nullstellensatz} cannot be strengthened to arbitrary
first-order (or first-order$^\star$ or even just coherent) formulas in place of
geometric$^\star$ sequents. For instance, in the case that~$\TT$ is the theory
of rings, the generic model~$U_\TT$ validates the coherent
formula~$\speak{any element~$x$ for which~$(x = 0 \Rightarrow 1 = 0)$ is invertible}$, but~$\ul{\TT}/U_\TT$
does not prove this fact, as it is for instance not validated by the polynomial
algebra~$U_\TT[X]$. However, Theorem~\ref{thm:nullstellensatz} still plays an
important role in understanding first-order formulas:

\begin{thm}\label{thm:characterization}
Let~$\TT$ be a geometric theory. Let~$\alpha$ be a first-order
formula over the signature of~$\TT$. Then the following statements are
equivalent.
\begin{enumerate}
\item The formula~$\alpha$ holds for~$U_\TT$. \smallskip
\item The formula~$\alpha$ is provable in first-order intuitionistic logic
modulo the axioms of~$\TT$ and the additional axioms
\begin{equation}\label{eq:nullstellensatz}\tag{$\ddagger$}
  \speak{$\sigma$ holds for~$U_\TT$} \Longrightarrow \speak{$\ul{\TT}/U_\TT$
  proves$^\star$~$\sigma$}
\end{equation}
where~$\sigma$ ranges over the geometric formulas over the signature of~$\TT$.
% XXX include free variables
\end{enumerate}
\end{thm}
